The title

### 2021-08-09 manuscript/manuscript.Rmd
New manuscript folder

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
# Load packages
library(papaja)
library(reticulate)
library(MASS)
library(Matrix)
library(lme4)
library(lmerTest)
library(buildmer)
library(emmeans)
library(magrittr)
library(tidyverse)
library(cowplot)

# Chunk options
knitr::opts_chunk$set(include = FALSE, echo = FALSE, fig.out = "100%", fig.width = 10)

# Conda environment for Python
#use_condaenv("r-reticulate", required = TRUE)
renv::use_python(python = "/Users/alexander/opt/anaconda3/bin/python", type = "conda")

# Bibliography file
r_refs("manuscript/r-references.bib")
```

```{r exp1-behavior}
# Function for processing behavioral data
preproc_behav <- function(txt_fname){
  # Read log file
  dat <- suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Rename some columns
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = as_factor(Wdh - min(Wdh) + 1)) %>%
    # Remove filler stimuli (i.e. well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  known <- dat %>% filter(part == "1" & Tastencode %in% c(201, 251)) %>% pull(item)
  insight <- dat %>% filter(part == "2" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>% pull(item)
  naive <- dat %>% filter(part == "2" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>% pull(item)
  # Create a new condition column, excluding trials with knwon stimuli or without condition
  dat %<>% mutate(condition = as_factor(case_when(item %in% known ~ "Excl_known",
                                                  item %in% insight ~ "Insight",
                                                  item %in% naive ~ "Naive",
                                                  TRUE ~ "Excl_cond")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if("richtig_inv" %in% dat$Bed){
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    dat %<>% select(part, condition, participant, item, RT)}
  return(dat)}

# List of filenames for behavioral data
fnames_exp1 <- list.files("data/exp1/RT", pattern = ".txt", full.names = TRUE)

# Apply the preprocessing function
data_exp1 <- map(fnames_exp1, preproc_behav)

# Tabulate percentage of stimuli per condition
(stims_exp1 <- bind_rows(data_exp1) %>% filter(condition != "Excl_cond") %>% 
    group_by(participant, condition) %>% tally() %>%
    group_by(condition) %>% summarise(mean = mean(n)/(60*3)) %>% # 60 items per condition
    mutate(mean = if_else(condition == "Excl_known", mean/2, mean))) # 120 items in total
```

```{python exp1-preprocessing}
# Import libraries
import mne
import glob
import os

# Define parameters for EEG preprocessing
preproc_params = dict(n_components=15, method='fastica',           # ICA parameters
                      l_freq=0.1, h_freq=30,                       # Filter edges
                      event_id={'match': 221, 'mismatch': 222},    # EEG triggers
                      tmin=-0.5, tmax=1.498,                       # Length of epochs
                      baseline=(-0.2,0),                           # Baseline correction
                      reject=dict(eeg=200e-6))                     # Rejection threshold

# # Debug
# locals().update(preproc_params)
# vhdr_fname = 'data/exp1/EEG/Vp0001.vhdr'
# metadata = metadata_exp1[0]

# Define preprocessing function
def preproc(vhdr_fname, metadata, n_components, method, l_freq, h_freq, event_id, tmin, tmax, baseline, reject):
  # Load EEG data
  raw = mne.io.read_raw_brainvision(vhdr_fname, preload=True)
  # Create new virtual EOG channels
  raw = mne.set_bipolar_reference(raw, 'Auge_u', 'Fp1', ch_name='VEOG', drop_refs=False)
  raw = mne.set_bipolar_reference(raw, 'F9', 'F10', ch_name='HEOG', drop_refs=False)
  raw.set_channel_types(mapping={'VEOG': 'eog', 'HEOG': 'eog'})
  # Add EasyCap electrode layout, removing any excessive channels
  montage = mne.channels.make_standard_montage('easycap-M1')
  raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(['VEOG', 'HEOG'])))
  raw.set_montage(montage=montage)
  # Re-reference to common average
  raw, _ = mne.set_eeg_reference(raw, 'average')
  # Run ICA on a copy of the data
  filt_raw = raw.copy()
  filt_raw.load_data().filter(l_freq=1, h_freq=None)
  ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
  ica.fit(filt_raw)
  # Remove bad components based on correlations with EOG
  eog_indices, eog_scores = ica.find_bads_eog(raw)
  ica.exclude = eog_indices
  raw = ica.apply(raw)
  # Apply band-pass filter
  raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
  # Epoching including baseline correction
  events, _ = mne.events_from_annotations(raw, verbose=False)
  epochs = mne.Epochs(raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True)
  # Add behavioral data
  epochs.metadata = metadata
  # Reject bad epochs
  epochs = epochs.drop_bad(reject=reject)

  # # Autoreject                    
  # import autoreject
  # 
  # ar = autoreject.AutoReject(thresh_method='random_search', verbose=False)
  # epochs = ar.fit_transform(epochs)
  # 
  # ransac = autoreject.Ransac(verbose=False)
  # epochs = ransac.fit_transform(epochs)
  # 
  # reject = autoreject.get_rejection_threshold(epochs)
  # epochs = epochs.drop_bad(reject=reject)

  return epochs

# List of raw EEG filenames
fnames_exp1 = sorted(glob.glob('data/exp1/EEG/*.vhdr'))

# Import behavioral data from R
metadata_exp1 = r.data_exp1

# Check if preprocessing was done already (delete the file to re-run)
if os.path.exists('output/exp1-epo.fif'):
  # Load preprocessed data from file
  epochs_exp1 = mne.read_epochs('output/exp1-epo.fif', preload=True)
else:
  # Apply the preprocessing function
  epochs_exp1 = [preproc(vhdr_fname=fname, metadata=meta, **preproc_params) for fname, meta in zip(fnames_exp1, metadata_exp1)]
  # Combine epochs into a single data set
  epochs_exp1 = mne.concatenate_epochs(epochs_exp1)
  # Backup epochs to the output folder
  epochs_exp1.save('output/exp1-epo.fif')

# # Get indices of bad epochs despite ICA
# rej_exp1 = epochs_exp1.drop_log
# rej_exp1[:] = [item for item in rej_exp1 if item != ['IGNORED']]
# rej_exp1 = [i for i in range(len(rej_exp1)) if rej_exp1[i] != []]

# Define function to compute grand-averaged ERPs per condition
def compute_evokeds(epochs):
  evokeds = dict()
  evokeds_dat = dict()
  # Each part becomes a dictionary of conditions
  for pt in ['1', '2', '3']:
    evokeds['Pt' + pt] = dict()
    evokeds_dat['Pt' + pt] = dict()
    # Each condition becomes a list of participants
    for cn in ['Insight', 'Naive']:
      evokeds['Pt' + pt][cn] = list()
      evokeds_dat['Pt' + pt][cn] = list()
      # For every participant we average trials separately for all parts and conditions
      for vp in epochs.metadata['participant'].unique():
        query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
        evokeds['Pt' + pt][cn].append(epochs[query].average())
      # Compute grand averages across participants for this part and condition
      evokeds['Pt' + pt][cn] = mne.grand_average(evokeds['Pt' + pt][cn])
      # Export only the actual data for R
      evokeds_dat['Pt' + pt][cn] = evokeds['Pt' + pt][cn].data
  return(evokeds, evokeds_dat)

# Apply the function to compute grand averages
evokeds_exp1, evokeds_dat_exp1 = compute_evokeds(epochs=epochs_exp1)

# Export electrode positions for R
elpos = epochs_exp1._get_channel_positions()

# # Plot evoked potentials
# mne.viz.plot_compare_evokeds(evokeds=evokeds_exp1['Pt1'], picks='P8'); plt.show()
# mne.viz.plot_evoked_topomap(evoked=evokeds_exp1['Pt2']['Insight']-evokeds_exp1['Pt2']['Naive'])
# evokeds_exp1['Pt2']['Insight'].data = evokeds_exp1['Pt2']['Insight'].data - evokeds_exp1['Pt2']['Naive'].data
# evokeds_exp1['Pt2']['Insight'].plot_topomap(times=[0.125, 0.175, 0.5], average=0.05, vmin=-1, vmax=1, cmap='viridis'); plt.show()
```

```{r exp1-analysis}
# Re-import behavioral data
data_exp1 <- py$epochs_exp1$metadata

# Factorize some columns
data_exp1 %<>% mutate(part = factor(part, levels = c("1", "2", "3")),
                      condition = factor(condition, levels = c("Insight", "Naive", "Excl_known", "Excl_cond")),
                      participant = factor(participant), item = factor(item))

# Check number of rejected epochs
rejected_exp1 <- data_exp1 %>% group_by(participant) %>% tally() %>% mutate(n = 360 - n) %>% pull(n)
mean(rejected_exp1); median(rejected_exp1); range(rejected_exp1)

# Create vectors with sample time points (in ms) and electrodes
tmin <- py$preproc_params$tmin*1000
tmax <- py$preproc_params$tmax*1000
sfreq <- py$epochs_exp1$info['sfreq']
times <- seq(tmin, tmax, 1000/sfreq)
els <- py$epochs_exp1$ch_names

# Define ERP components of interest
comps <- tibble(name = c("P1", "N1", "N400"),
                start = c(100, 150, 400),
                stop = c(150, 200, 700),
                roi = list(c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
                           c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
                           c("C1", "C2", "Cz", "CP1", "CP2", "CPz")))

# Define function to compute single-trial ERPs
compute_erps <- function(epochs, els, name, start, stop, roi){
  erps <- epochs[, which(els %in% roi), which(times %in% start:stop)]
  erps <- apply(erps, 1, mean, na.rm = TRUE)
  return(erps)}

# Check if ERPs were computed already (delete the file to re-run)
if (file.exists("output/exp1-erps.RDS")){
  # Load ERPs from file
  data_exp1 <- readRDS("output/exp1-erps.RDS")
} else {
  # Import epochs from Python, converting Volts to Microvolts
  epochs_exp1 <- py$epochs_exp1$get_data()*1e6
  # Apply the ERP function and save to output folder
  data_exp1 <- pmap_dfc(comps, compute_erps, epochs = epochs_exp1, els = els) %>%
    set_names(comps$name) %>% cbind(data_exp1, .) %T>% saveRDS("output/exp1-erps.RDS")}

# Remove any trials that don't beling to either of the two conditions
data_exp1 %<>% filter(condition %in% c("Insight", "Naive")) %>% droplevels()

# Sliding difference coding for condition (insight-naive) and part (2-1, 3-1)
t(contrasts_condition <- t(cbind(c("Insight" = 1, "Naive" = -1))))
t(contrasts_part <- t(cbind(c("1" = -1, "2" = 1, "3" = 0),
                            c("1" = 0, "2" = -1, "3" = 1))))
contrasts(data_exp1$condition) <- ginv(contrasts_condition)
contrasts(data_exp1$part) <- ginv(contrasts_part)

# Formula and parameters for linear mixed-effects regression models (LMMs)
form_exp12 <- tabulate.formula(~ part*condition + (part*condition|participant) + (part*condition|item))
form_exp12 %<>% mutate(block = replace(block, is.na(grouping), "fixed"))
ctrl_params <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Formula for follow-up contrasts
specs_exp12 = pairwise ~ condition|part

# Define function to compute LMMs and contrasts for each component
compute_models <- function(dep, formula, data, control, specs){
  browser()
  build <- buildmer(dep = dep, formula = formula, data = data, ddf = "Satterthwaite", calc.anova = TRUE, control = control)
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  conts <- emmeans(build@model, specs = specs, infer = TRUE, data = data)$contrasts %>% as.data.frame()
  mod <- list("model" = build@model, "anova" = build@anova, "summary" = build@summary, "contrasts" = conts)
  return(mod)}

# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp1-stats.RDS")){
  # Load models from file
  models_exp1 <- readRDS("output/exp1-stats.RDS")
} else {
  # Apply the modelling function and save to output folder
  models_exp1 <- map(comps$name, compute_models, formula = form_exp12, data = data_exp1, control = ctrl_params,
                     specs = specs_exp12) %>% set_names(comps$name) %T>% saveRDS("output/exp1-stats.RDS")}

# Store significance levels for plotting
stars_exp1 = list(P1 = c(Pt1 = NA, Pt2 = NA, Pt3 = "*"),
                  N1 = c(Pt1 = NA, Pt2 = "**", Pt3 = NA),
                  N400 = c(Pt1 = NA, Pt2 = "***", Pt3 = "**"))
```

```{r exp1-table, include=TRUE, results="asis"}
# Check if models were computed already (delete the file to re-run)
if (file.exists("output/exp1-stats.RDS")){
  # Load models from file
  models_exp1 <- readRDS("output/exp1-stats.RDS")
} else {
  # Apply the modelling function and save to output folder
  models_exp1 <- map(comps$name, compute_models, formula = form_exp12, data = data_exp1, control = ctrl_params,
                     specs = specs_exp12) %>% set_names(comps$name) %>% saveRDS("output/exp1-stats.RDS")}

# Define function to create ANOVA-style table
create_table <- function(models, stub.anova, stub.contrasts, caption, note){
  anov <- map(models, function(model){
    data.frame("stat" = paste0(format(round(model$anova$`F value`, 2), trim = FALSE, nsmall = 2), " (", model$anova$NumDF,
                               ", ", format(round(model$anova$DenDF, 1), trim = FALSE, nsmall = 1), ")"),
               "p" = format(round(model$anova$`Pr(>F)`, 3), trim = FALSE, nsmall = 3)) %>%
      mutate(p = ifelse(p == "0.000", "<.001", substr(p, 2, nchar(p))))}) %>%
    bind_cols() %>% set_rownames(c(stub.anova))
  cnames <- c("\\textit{Type III}", rep(c("\\textit{F} (\\textit{df})", "p" = "\\textit{p}"), 3))
  conts <- map(models, function(model){
    data.frame("stat" = paste0(format(round(model$contrasts$estimate, 2), trim = FALSE, nsmall = 2), " [",
                               format(round(model$contrasts$lower.CL, 2), trim = FALSE, nsmall = 2), ", ",
                               format(round(model$contrasts$upper.CL, 2), trim = FALSE, nsmall = 2), "]"),
               "p" = format(round(model$contrasts$`p.value`, 3), trim = FALSE, nsmall = 3)) %>%
      mutate(p = ifelse(p == "0.000", "<.001", substr(p, 2, nchar(p)))) %>%
      add_row("stat" = "Est. [95% CI]", "p" = "\\textit{p}", .before = 1)}) %>% 
    bind_cols() %>% set_rownames(c("\\textit{Insight $-$  naive}", stub.contrasts))
  apa_table(list(anov, conts), col.names = cnames, escape = FALSE,
            col_spanners = list("\\textbf{P1}" = 2:3, "\\textbf{N1}" = 4:5, "\\textbf{N400}" = 6:7),
            midrules = c(nrow(anov) + 1), font_size = "footnotesize", align = "lcccccc",
            caption = caption, note = note)}

>>>>

---
title             : "Instant effects of semantic information on visual perception"
shorttitle        : "Semantically informed perception"
author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : yes
    address       : "Rudower Chaussee 18, 12489 Berlin, Germany"
    email         : "alexander.enge@hu-berlin.de"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands"
authornote: |
  \addORCIDlink{Alexander Enge}{0000-0003-0100-2297}

  \addORCIDlink{Rasha Abdel Rahman}{0000-0002-8438-1570}

  The authors declare no competing financial interests.
  The preprocessed data and analaysis code for this study are openly available at https://osf.io/uksbc/.
abstract: |
  Does our perception of an object change as soon as we discover what function it serves?
  This question matters not only when we encounter novel tools or gadgets in our everyday lives, but also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities.
  Here, we showed human participants (*n* = 48) pictures of unfamiliar objects either together with words matching their function, leading to semantically informed perception, or together with non-matching words, resulting in naive perception.
  We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
  We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component (150--200 ms) and reduced amplitudes in the N400 component (400--700 ms).
  When the same objects were presented once more without any information, the N400 effect persisted and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
  Consistent with previous work, our results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
  We demonstrate for the first time that these effects are instantaneous in the sense that they can be observed within the same trial as well as one trial after the information has been given, providing strong evidence that cognition affects perception.
keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
bibliography      : ["references.bib", "r-references.bib"]
csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man"
fontsize          : "10pt"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
  papaja::apa6_docx: default
editor_options: 
  chunk_output_type: console
header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load packages
library(buildmer)
library(cowplot)
library(emmeans)
library(here)
library(knitr)
library(lme4)
library(lmerTest)
library(magick)
library(MASS)
library(papaja)
library(parallel)
library(reticulate)
library(scales)
library(tidyverse)
library(magrittr)

# Load helper functions
source(here("manuscript", "plots_tables.R"))

# Write R package citations
r_refs(here("manuscript", "r-references.bib"), append = FALSE)

# Set global chunk options
opts_chunk$set(
  fig.align = "center",
  fig.width = 10,
  include = FALSE,
  out.width = "100%"
)

# Re-run preprocessing and/or model fitting?
run <- list(preprocessing = FALSE, model_fitting = FALSE)

# Number of CPU cores for parallel model fitting
n_cores <- detectCores()

# File names for easier I/O
deps <- c("logRT", "P1", "N170", "N400")
fnames <- list(
  montage = "results/montage.csv",
  obj_per_cond = "results/obj-per-cond.csv",
  trials = "results/trials.csv",
  ave_grand = "results/ave_grand.csv",
  ave_tfr_flat = "results/ave-tfr_flat.csv",
  ave_tfr_diff = "results/ave-tfr_diff.csv",
  models_lme = "results/models_lme.RDS",
  models_brm = str_c("results/model_brm_", deps, ".RDS") %>% set_names(deps),
  model_brm_prior = "results/model_brm_prior.RDS",
  models_control = "results/models_control.RDS"
)
```

# Introduction

Does our perception of an object change as soon as we discover what function it serves?
This question speaks not only to our everyday experiences, where we encounter novel tools and gadgets in our dynamic working and private environments.
It also pertains to the long-standing debate about the cognitive (im)penetrability [@pylyshyn1999] of perception by higher-level capacities such as semantic knowledge or language.
According to one view, these cognitive capacities kick in only after the retinal input has been processed by a specialized module for visual perception [@firestone2016; @fodor1983].
This module is encapsulated from higher-level inputs and therefore processes the visual information in a purely feedforward fashion, progressing from lower areas with small receptive field sizes to areas representing increasingly complex shapes and, eventually, whole objects [@dicarlo2012].
This architecture has been mirrored in classical [e.g., @marr1982] and contemporary [e.g., @krizhevsky2012] computer vision models, allowing them to achieve human-level perfomance in predicting object category labels from images [for review, see @lindsay2020].

The cognitive impenetrability hypothesis is challenged by the alternative view that perception dynamically interacts with different aspects of cognition from early on [@churchland1994; @lupyan2015].
This view is supported by a variety of theoretical, behavioral, and neurophysiological accounts.

On the theoretical level, the reverse hierarchy theory [@ahissar2004; @hochstein2002] posits that conscious processing initally occurs at the level of whole objects or object categories.
Only after this high-level interpretation has been obtained, more fine grained visual details---if relevant for the current task---are being accessed via top-down connections.
Along similar lines, Bayesian inference and predictive coding theories [e.g., @clark2013; @panichello2013; @yuille2006] propose bi-directional loops of predictions and prediction errors between visual and non-visual processes.

On the behavioral level, these ideas are supported by psychological studies that showed differences in ratings, detection rates, or reaction times for visual stimuli depending on their emotional [e.g., @phelps2016], motivational [e.g., @balcetis2010], linguistic [e.g., @boutonnet2015] or semantic [e.g., @gauthier2003] background.
However, some of these studies confounded tentantive high-level effects with low-level differences between conditions.
Furthermore, many studies were not able to discern between perceptual and post-perceptual (e.g., memory-related) effects [@firestone2016].

On the neurophysiological level, event-related potentials (ERPs) measured from the human EEG can mitigate most of these concerns:
Their excellent temporal resolution allows us to probe how early we can detect influences of high-level (e.g., semantic) information.
To this end, participants were typically trained to associate visual objects with different amounts of semantic information.
After training, an orthogonal task was used to compare the ERPs in response to semantically trained objects versus non-semantically trained objects.
This revealed differences not only in late ERP components associated with semantic processing (i.e., the N400 component) but also in the visual P1 component [@abdelrahman2008; @maier2019; @samaha2018; @weller2019].
The early peak of this component and its source in the occipital cortex [@abdelrahman2008] point to an immediate effect of semantic knowledge on visual perception.

Here we measured ERPs in response to unfamiliar objects instantly while participants gained an semantic understanding of their function.
To this end, we presented half of the objects together with matching keywords, thus allowing participants to understand what kind of object they were viweing, or together with non-matching keywords, thus keeping the perception of the object semantically naive.
We then presented the same objects again to test for downstream effects of semantic information as in previous studies.
We examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).

# Materials and Methods

## Participants

Participants were 48 German native speakers (31 female, 17 male) with a mean age of 23.5 years (range 18--32) and no history of psychological disorder or treatment.
No a priori power analysis was carried out and the sample size was chosen in line with previous EEG studies on object processing in our lab.
All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision.
They provided written informed consent before starting the experiment and received a compensation of €8 per hour for participating.

## Materials

Stimuli consisted of 240 grayscale photographs of real-world objects.
Of these, 120 were well-known everyday objects (e.g., a bicycle, a toothbrush) and served as filler stimuli of no interest.
The other 120 were rare objects presumed to be unfamiliar to the majority of participants (e.g., a galvanometer, an udu drum; see Appendix A).
All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz.
At a standardized viewing distance of 90 cm, the images subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

For each unfamiliar object, a pair of German keywords---a noun and a verb---was selected, describing the typical function or use of the object in a way that could typically be related to its visual features and their configuration (e.g., voltage, measuring; clay pot, drumming).
As our central experimental manipulation, the presentation of half of the objects was preceded by keywords that correctly matched their respective function, whereas the presentation of the other half of the objects was preceded by non-matching keywords belonging to one of the other objects.
The matching keywords were expected to induce semantically informed perception (i.e., participants suddenly understanding what kind of object they were viewing), whereas the non-matching keywords were expected to prevent such an understanding and keep the perception of the object semantically naive.
All participants saw each unfamiliar object with only one type of keywords (matching or non-matching).
This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to naive perception) to an equal number of participants.
The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

## Experimental Design

The experiment consisted of three parts (see Figure \@ref(fig:fig1)A).
In the *pre-insight* part, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords.
Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s.
The inter-trial interval was 0.5 s and participants took a self-timed break after each block of 60 objects.
The task, which was kept the same across all three parts, was to classify each object using one of four response alternatives: (a) "I know what this is or have a strong assumption," (b) "I have an assumption what this is," (c) "I have rather no assumption what this is," or (d) "I don't know what this is and have no assumption." Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively.
The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

```{r, preprocessing}
# List EEG files and behavioral log files
fname_vhdr <- c(
  list.files("data/raw/eeg/exp1", pattern = ".vhdr", full.names = TRUE),
  list.files("data/raw/eeg/exp2", pattern = ".vhdr", full.names = TRUE)
)
fname_log <- c(
  list.files("data/raw/rt/exp1", pattern = ".txt", full.names = TRUE),
  list.files("data/raw/rt/exp2", pattern = ".txt", full.names = TRUE)
)

# Generate new BIDS-style subject IDs
seq(length(fname_vhdr)) %>%
  str_pad(width = 2, pad = 0) %>%
  str_c("sub-", .) -> subject_id

# Define ERP components of interest
erp_components <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(0.1, 0.15, 0.4),
  tmax = c(0.15, 0.2, 0.7),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

# Define frequencies and number of cycles for TFR analysis
tfr_freqs <- seq(4, 50, by = 2)
tfr_cycles <- seq(3, 10, length.out = length(tfr_freqs))

# Run preprocessing for all subjects using MNE-Python (see preprocessing.py)
# Will run only if `run$preprocessing = TRUE` and if raw data are available
if (run$preprocessing) {

  # Import the custom Python functions for preprocessing
  pp <- import_from_path("preprocessing", path = "code")

  # Do the actual preprocessing
  tibble(fname_vhdr, fname_log, subject_id) %>%
    pwalk(
      pp$preprocess,
      output_dir = "data/preprocessed",
      eog_names = c("HEOG", "VEOG"),
      eog_anodes = c("F9", "Auge_u"),
      eog_cathodes = c("F10", "Fp1"),
      montage_kind = "easycap-M1",
      ica_components = as.integer(15),
      ica_method = "fastica",
      ica_seed = as.integer(1234),
      hipass = 0.1,
      lowpass = 30,
      event_id = dict(match = as.integer(221), mismatch = as.integer(222)),
      tmin = -0.5,
      tmax = 1.489,
      baseline = tuple(-0.2, 0),
      reject = dict(eeg = 200e-6),
      erp_components = erp_components,
      average_by = c("part", "condition", "subject_id"),
      tfr_baseline = tuple(-0.3, -0.1),
      tfr_equalize = "condition == ['Informed', 'Naive']",
      tfr_resample = py_none(),
      tfr_freqs = tfr_freqs,
      tfr_cycles = tfr_cycles,
      tfr_crop = tuple(-0.2, 0.8)
    )

  # Combine single trial data for all subjects
  dir.create("results")
  str_c("data/preprocessed/", subject_id, "_trials.csv") %>%
    map(read_csv, col_types = cols()) %>%
    bind_rows() %>%
    write_csv(fnames$trials)

  # Get standard EasyCap electrode positions and project them into 2D space
  read_csv("data/preprocessed/sub-01_tfr-ave.csv", col_types = cols()) %>%
    select(-c(part, condition, subject_id, time, freq)) %>%
    colnames() -> ch_names
  import("mne.channels._standard_montage_utils")$MONTAGE_PATH %>%
    str_c("/easycap-M1.txt") %>%
    read_tsv(col_types = cols()) %>%
    rename(electrode = Site) %>%
    inner_join(tibble(electrode = ch_names), ., by = "electrode") %>%
    mutate(
      x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
      y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi)),
    ) %>%
    mutate(across(c(x, y), round, 3)) %>%
    write_csv(fnames$montage)

  # Determine number of objects per condition for each participant
  fname_log %>%
    map2(subject_id, pp$read_log) %>%
    bind_rows() %>%
    with(table(subject_id, condition) / 3) %>%
    as.data.frame.matrix() %>%
    rownames_to_column("subject_id") %>%
    write_csv(fnames$obj_per_cond)
}
```

```{python, grand_averaging, eval=run$preprocessing}
import numpy as np
import pandas as pd

# Read evoked *potentials* for all subjects
subs = r.subject_id
fnames_ave = [f"data/preprocessed/{sub}_ave.csv" for sub in subs]
ave = [pd.read_csv(fname) for fname in fnames_ave]
ave = pd.concat(ave)

# Average across subjects
id_vars = ["part", "condition", "time"]
ave = ave.groupby(id_vars, as_index=False).mean()
ave.to_csv(r.fnames["ave_grand"], float_format="%.3f", index=False)

# Read evoked *power* for all subjects
fnames_ave_tfr = [f"data/preprocessed/{sub}_tfr-ave.csv" for sub in subs]
ave_tfr = [pd.read_csv(fname) for fname in fnames_ave_tfr]
ave_tfr = pd.concat(ave_tfr)

# # Subject 18 is weird!
# ave_tfr = ave_tfr[ave_tfr["subject_id"] != "sub-18"]

# Compute difference in power between conditions
id_vars = ["part", "condition", "subject_id", "time", "freq"]
electrodes = [col for col in ave_tfr.columns if (col not in id_vars)]
ave_tfr_diff = (
    ave_tfr.query("condition == 'Informed'")[electrodes]
    - ave_tfr.query("condition == 'Naive'")[electrodes].values
)
ave_tfr_diff[id_vars] = ave_tfr.query("condition == 'Informed'")[id_vars]

# Compute flat average across conditions
id_vars.remove("condition")
ave_tfr_flat = ave_tfr.groupby(id_vars, as_index=False).mean()

# Average across subjects and save
id_vars.remove("subject_id")
dfs = [ave_tfr_flat, ave_tfr_diff]
fnames_tfr = [r.fnames["ave_tfr_flat"], r.fnames["ave_tfr_diff"]]
for df, fname in zip(dfs, fnames_tfr):
    df = df.groupby(id_vars, as_index=False).mean()
    df.to_csv(fname, float_format="%.3f", index=False)
del df, dfs, ave, ave_tfr, ave_tfr_diff, ave_tfr_flat
```

```{r, models_lme}
# Compute average number of objects per condition
(obj_per_cond <- read_csv(fnames$obj_per_cond, col_types = cols()) %>%
  summarise(across(.cols = -subject_id, .fns = mean)) %>%
  as.list())

# Read single trial data for all subjects
str_c("data/preprocessed/", subject_id, "_trials.csv") %>%
  map(read_csv, col_types = cols()) %>%
  bind_rows() %>%
  mutate(
    part = factor(part, levels = c("I", "II", "III")),
    subject_id = factor(subject_id),
    item_id = factor(item_id),
    RT = ifelse(RT < 200, NA, RT),
    logRT = log(RT)
  ) -> trials

# Record number of rejected ERP epochs per participant
(rejected <- trials %>%
  count(subject_id) %>%
  mutate(n = 360 - n) %>%
  pull(n))

# Remove trials with unrealistically short RTs or invalid conditions
sum(is.na(trials$RT))
sum(is.na(trials$RT)) / nrow(trials)
trials %<>% na.omit()

# Contrast coding for parts
cbind(c("I" = -1, "II" = 1, "III" = 0), c("I" = 0, "II" = -1, "III" = 1)) %>%
  t() %>%
  ginv() -> contrasts(trials$part)
